---
title: "Time-Series Forecasting of Banana Prices"
author: "Kaisa Roggeveen, Scott Graham, Irwin Khuu, Johnson Tran, David Lin"
ioslides_presentation:
  smaller: true
  logo: Images/banana.png
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(knitr, warn.conflicts = FALSE, quietly = FALSE)
library(pander, warn.conflicts = FALSE, quietly = FALSE)
library(forecast, warn.conflicts = FALSE, quietly = TRUE)
library(tseries, warn.conflicts = FALSE, quietly = TRUE)
library(tidyverse, warn.conflicts = FALSE, quietly = TRUE)
library(lubridate, warn.conflicts = FALSE, quietly = TRUE)
library(tidyquant, warn.conflicts = FALSE, quietly = TRUE)
library(gridExtra, warn.conflicts = FALSE, quietly = TRUE)

source("ggacf.R", echo = FALSE)
source("ggcor.R", echo = FALSE)

log_ret_to_price <- function(historical_data, forecast_data){
  historical_data %>% 
    bind_rows(
      bind_cols(
        Date = 
          historical_data %>% 
          filter(
            Date == max(Date)
            ,Type == "Forecast"
          ) %>% 
          .$Date + months(1)
        ,(
          historical_data %>% 
            filter(
              Date == max(Date)
              ,Type == "Forecast"
            ) %>%
            select(-Date, -Type) %>% 
            mutate_all(funs(./.*Mean))
        ) * 
          (
            forecast_data %>% 
              filter(
                Date == max(historical_data$Date) + months(1)
              ) %>% 
              select(-Date, -Type) %>%
              exp()
          )
        ,Type = "Forecast"
      )
    )
}

num_months_name <- 
  tibble(
    Name =
      as.ordered(
        c(
          "Jan"
          ,"Feb"
          ,"Mar"
          ,"Apr"
          ,"May"
          ,"Jun"
          ,"Jul"
          ,"Aug"
          ,"Sep"
          ,"Oct"
          ,"Nov"
          ,"Dec"
        )
      )
    ,Num = 1:12
    ,Season = 
      as.ordered(
        c(
          rep("Winter", each = 2)
          ,rep(c("Spring", "Summer", "Fall"), each = 3)
          ,"Winter"
        )
      )
  )
num_months_name$Name <- factor(num_months_name$Name, levels = num_months_name$Name)
num_months_name$Season <- factor(num_months_name$Season, levels = c("Spring", "Summer", "Fall", "Winter"))

theme_minimal2 <- theme_minimal() %>%  theme_set()
theme_minimal2 <-
  theme_update(
    panel.border = element_rect(
      linetype = "solid"
      ,colour = "grey92"
      ,fill = NA
    )
    ,strip.background = element_rect(
      linetype = "solid"
      ,colour = "grey92"
      ,fill = NA
    )
  )

# Data Setup and Retieval ----------
# start_date <- as_date("2018-03-13")
end_date <- as_date("2017-01-01")

# Anything of the form: https://fred.stlouisfed.org/series/*, where * is the ticker that is looked up in the first parameter of tq_get
# https://fred.stlouisfed.org/series/DPROPANEMBTX Mount Belvieu Propane Prices
commodity_prices <- 
  tq_get(
    x = tibble(commodity = "PBANSOPUSDM")
    ,get = "economic.data"
    ,from = end_date - years(20)
    ,to = end_date
  ) %>%
  filter(!is.na(price)) %>% 
  group_by(commodity) %>% 
  tq_mutate(
    select = price
    ,mutate_fun = periodReturn
    ,period = "daily"
    ,col_rename = "R_a"
  )

banana_price <-
  commodity_prices %>% 
  filter(commodity == "PBANSOPUSDM") %>% 
  ungroup() %>% 
  select(-commodity) %>% 
  rename(
    Date = date
    ,Price = price
  ) %>% 
  mutate(
    Year = year(Date)
    ,Month = month(Date)
    ,Day = day(Date)
  ) %>% 
  left_join(
    num_months_name
    ,by = c("Month" = "Num")
  ) %>% 
  rename(`Month Name` = Name)

num_periods_ahead <- 12*3
```
\[
\newcommand{\Prob}{\operatorname{P}}
\newcommand{\E}{\operatorname{E}}
\newcommand{\Var}{\operatorname{Var}}
\newcommand{\Cov}{\operatorname{Cov}}
\newcommand{\se}{\operatorname{se}}
\newcommand{\re}{\operatorname{re}}
\newcommand{\ybar}{{\overline{Y}}}
\newcommand{\phat}{{\hat{p}}}
\newcommand{\that}{{\hat{T}}}
\newcommand{\med}{{\tilde{Y}}}
\newcommand{\Logit}{{\operatorname{Logit}}}
\]

## Table of Contents
| Section                  | Slide |
|--------------------------|------:|
| Methodology              |     4 |
| Model Selection          |    14 |
| Forecasting Banana Price |    18 |
| Model Diagnostics        |    23 |
| Results                  |    26 |


## Introduction
- Collected from the Federal Reserve Bank of St. Louis

- Bananas are typically consumed fresh

- Look for Seasonality

- Look for Trend


# Methodology
## Original Time Series Plot
```{r Original TS, fig.height=3, fig.width=7}
banana_price %>%
  ggplot(
    aes(
      x = Date
      ,y = Price
    )
  ) +
  geom_smooth(method = "loess", se = FALSE, colour = "#66c2a5") +
  geom_smooth(method = "lm", se = FALSE, colour = "#fc8d62") +
  geom_line() +
  labs(
    title = "Banana Price"
    ,y = "Price (USD/Metric Ton)"
  )
```


## Original Seasonal Plot
```{r Price Violin Plot}
banana_price %>%
  ggplot(
    aes(
      x = Season
      ,y = Price
    )
  ) +
  geom_violin(fill = NA) +
  geom_boxplot(fill = NA, width = 0.1, outlier.colour = NA) +
  geom_jitter(aes(colour = Year)) +
  facet_wrap(
    ~ Season
    ,nrow = 1
    ,scales = "free_x"
  ) +
  scale_color_distiller(
    type = "seq"
    ,palette = "RdYlBu"
  ) +
  labs(
    title = "Seasonal Banana Price Violin Plot"
    ,y = "Price (USD/Metric Ton)"
  ) +
  theme(
    axis.text.x = element_blank()
    ,axis.title.x = element_blank()
  )
```


## Lagged Scatter Plot
```{r Lagged Scatter Plot, fig.height=5, fig.width=5}
banana_lagged <-
  banana_price %>%
  mutate(
    `Lag 01` = lag(Price, n = 1)
    ,`Lag 02` = lag(Price, n = 2)
    ,`Lag 03` = lag(Price, n = 3)
    ,`Lag 04` = lag(Price, n = 4)
    ,`Lag 05` = lag(Price, n = 5)
    ,`Lag 06` = lag(Price, n = 6)
    ,`Lag 07` = lag(Price, n = 7)
    ,`Lag 08` = lag(Price, n = 8)
    ,`Lag 09` = lag(Price, n = 9)
    ,`Lag 10` = lag(Price, n = 10)
    ,`Lag 11` = lag(Price, n = 11)
    ,`Lag 12` = lag(Price, n = 12)
  )

banana_lagged %>%
  gather(
    key = Lag
    ,value = `Lagged Price`
    ,-Date
    ,-Year
    ,-Month
    ,-`Month Name`
    ,-Season
    ,-Day
    ,-Price
    ,-R_a
  ) %>%
  filter(!is.na(`Lagged Price`)) %>%
  ggplot(
    aes(
      x = Price
      ,y = `Lagged Price`
    )
  ) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "loess", se = FALSE, colour = "#66c2a5") +
  geom_smooth(method = "lm", se = FALSE, colour = "#fc8d62") +
  facet_wrap(~ Lag) +
  labs(
    title = "Lag Scatter Plot"
    ,x = "Price (USD/Metric Ton)"
    ,y = "Lagged Price (USD/Metric Ton)"
  )
```


## Autocorrelation Function
```{r Price ACF, fig.height=4,fig.width=7}
ggacf(banana_price, col = Price, type = "correlation") + labs(title = "Price ACF")
```


## Partial Autocorrelation Function
```{r Price PACF, fig.height=4, fig.width=7}
ggacf(banana_price, col = Price, type = "partial") + labs(title = "Price PACF")
```


## Transformation Time-Series Plot
```{r Trans TS Plot, fig.height=6,fig.width=7}
banana_price <-
  banana_price %>%
  mutate(
    `Log Price` = log(Price)
    ,`Log Return` = log(1 + R_a)
    ,`Box Cox Price` = BoxCox(Price, lambda = BoxCox.lambda(Price))
    ,`Box Cox Return` = BoxCox(Price, lambda = BoxCox.lambda(R_a))
  )

banana_price %>%
  gather(
    key = Transformation
    ,value = Value
    ,-Date
    ,-Year
    ,-Month
    ,-`Month Name`
    ,-Season
    ,-Day
  ) %>%
  ggplot(
    aes(
      x = Date
      ,y = Value
    )
  ) +
  geom_smooth(method = "loess", se = FALSE, colour = "#66c2a5") +
  geom_smooth(method = "lm", se = FALSE, colour = "#fc8d62") +
  geom_line() +
  facet_wrap(
    ~ Transformation
    ,scales = "free_y"
    ,ncol = 1
  ) +
  labs(title = "Time Series Plot of Transformations")
```



## Transformation Autocovariance Function
```{r Return AcF and PACF}
banana_R_a_acf <- ggacf(banana_price, col = R_a, type = "correlation") + labs(title = "R_a ACF")
banana_R_a_pacf <- ggacf(banana_price, col = R_a, type = "partial") + labs(title = "R_a PACF")
banana_log_ret_acf <- ggacf(banana_price, col = `Log Return`, type = "correlation") + labs(title = "Log Return ACF")
banana_log_ret_pacf <- ggacf(banana_price, col = `Log Return`, type = "partial") + labs(title = "Log Return PACF")
grid.arrange(banana_R_a_acf, banana_R_a_pacf, banana_log_ret_acf, banana_log_ret_pacf, top = "ACF & PACF Plots for Return on Bananas")
```


## Decomposition of Transformation
```{r Decomposition Stats}
banana_logreturn_decomp <-
  banana_price$`Log Return` %>%
  ts(frequency = 12) %>%
  stl(s.window = "periodic")
```

```{r Log Return Decomposition Plot, fig.height=5, fig.width=7}
banana_price %>%
  select(
    Date
    ,Year
    ,Month
    ,`Month Name`
    ,Day
    ,`Log Return`
  ) %>%
  bind_cols(
    banana_logreturn_decomp$time.series %>%
      data.frame() %>%
      as_tibble() %>%
      rename(
        Remainder = remainder
        ,Seasonal = seasonal
        ,Trend = trend
      )
  ) %>%
  rename(Data = `Log Return`) %>%
  gather(
    key = Decomposition
    ,value = `Log Return`
    ,-Date
    ,-Year
    ,-Month
    ,-`Month Name`
    ,-Day
  ) %>%
  mutate(
    Decomposition = factor(Decomposition, levels = c("Data", "Trend", "Seasonal", "Remainder"))
  ) %>%
  ggplot(
    aes(
      x = Date
      ,y = `Log Return`
    )
  ) +
  geom_line() +
  facet_wrap(
    ~ Decomposition
    ,ncol = 1
    ,scales = "free_y"
  ) +
  labs(title = "Banana Log Returns Decomposition")
```


## ADF Test (Dickey's Test)
```{r ADF}
suppressWarnings(
  banana_price %>%
  {tseries::adf.test(.$Price, alternative = "stationary")}
)

suppressWarnings(
  banana_price %>%
  {tseries::adf.test(.$`Log Return`, alternative = "stationary")} 
)
```


# Model Selection 
## Auto ARIMA
```{r Auto-ARIMA}
banana_log_ret_autoarima_01 <- auto.arima(banana_price$`Log Return`, seasonal = TRUE)
# summary(banana_log_ret_autoarima_01)
```
$$\phi(B)\Phi(B^s)Y_t= \theta(B)\Theta(B^s)Z_t $$ 
where 
$$\phi(z) = 1-\phi_1(z)-...-\phi_pz^p,\Phi(z) = 1-\Phi_1(z)-...-\Phi_pz^p $$
$$\theta(z) = 1+\theta_1z+...+\theta_qz^q, \Theta(z) = 1+\Theta_1z+...+\Theta_qz^q $$
Auto-ARIMA(1,0,1) where 
$$\phi(z) = 1-0.7300z$$
$$\theta(z) = 1-0.9256z$$


## Minimum AIC
```{r min AIC BIC, cache=TRUE}
pdqPDQ <-
  0:1 %>% 
  rep(times = 4, each = 3) %>% 
  combn(m = 6) %>% 
  t() %>% 
  unique()

banana_arima_builder <- function(pdq){
  banana_arima <-
    Arima(
      y = banana_price$`Log Return`
      ,order = pdq[1:3]
      ,seasonal = pdq[4:6]
    )
  banana_aic <- banana_arima$aic
  banana_bic <- AIC(banana_arima, k = as.numeric(count(banana_price)))
  list(
    p = pdq[1]
    ,d = pdq[2]
    ,q = pdq[3]
    ,P = pdq[4]
    ,D = pdq[5]
    ,Q = pdq[6]
    # ,arima = banana_arima
    ,AIC = banana_aic
    ,BIC = banana_bic
  )
}

banana_arima_stats <- 
  1:dim(pdqPDQ)[1] %>% 
  map(~ pdqPDQ[., ]) %>% 
  map_df(banana_arima_builder)
```

```{r min aic}
banana_min_aic <- 
  banana_arima_stats %>% 
  filter(AIC == min(AIC)) %>% 
  unique()
banana_min_aic %>% 
  kable(caption = "Minimum AIC")

banana_aic_arima <- 
  banana_min_aic %>% 
  select(-AIC, -BIC) %>%
  mutate(`Model #` = 1:nrow(banana_min_aic)) %>%
  gather(
    key = Letter
    ,value = Num
    ,-`Model #`
  ) %>% 
  split(.$`Model #`) %>% 
  map(~ .$Num) %>% 
  map(
    ~ Arima(
      y = banana_price$`Log Return`
      ,order = .[1:3]
      ,seasonal = .[4:6]
    )
  )
# banana_aic_arima %>% 
#   map(summary)

# Just 1 entry
# banana_aic_arima$`1` %>% 
#   summary() %>% 
#   pander()
```


## Minimum BIC
```{r min bic}
banana_min_bic <- 
  banana_arima_stats %>% 
  filter(BIC == min(BIC)) %>% 
  unique()
banana_min_bic %>% 
  kable(caption = "Minimum BIC")

banana_bic_arima <- 
  banana_min_bic %>% 
  select(-AIC, -BIC) %>%
  mutate(`Model #` = 1:nrow(banana_min_bic)) %>%
  gather(
    key = Letter
    ,value = Num
    ,-`Model #`
  ) %>% 
  split(.$`Model #`) %>% 
  map(~ .$Num) %>% 
  map(
    ~ Arima(
      y = banana_price$`Log Return`
      ,order = .[1:3]
      ,seasonal = .[4:6]
    )
  )
# banana_bic_arima %>% 
#   map(summary)

# Just 1 entry
# banana_bic_arima$`1` %>% 
#   summary() %>% 
#   pander()
```

$$\phi(B)\Phi(B^s)Y_t= \theta(B)\Theta(B^s)Z_t $$ where $$\phi(z) = 1-\phi_1(z)-...-\phi_pz^p,\Phi(z) = 1-\Phi_1(z)-...-\Phi_pz^p $$
$$\theta(z) = 1+\theta_1z+...+\theta_qz^q, \Theta(z) = 1+\Theta_1z+...+\Theta_qz^q $$
AIC (1,0,1)
where $$\phi(z) = 1-0.7573z$$ $$\theta(z) = 1-0.9568z $$

BIC (0,1,0)(0,1,1)
where $$\phi(z) = 1-0.7573z$$ $$\theta(z) = 1-0.9568z$$

# Forecasting Banana Price
```{r AA Forecast Log Return, fig.height=3, fid.width=7}
banana_autoarima_forecast <- 
  banana_log_ret_autoarima_01 %>% 
  forecast(num_periods_ahead) 

banana_autoarima_forecast <-
  tibble(
    Date = max(banana_price$Date) + months(1:num_periods_ahead)
    ,Mean = as.numeric(banana_autoarima_forecast$mean)
  ) %>% 
  bind_cols(as_tibble(banana_autoarima_forecast$lower)) %>% 
  rename(`Lower 80%` = `80%`, `Lower 95%` = `95%`) %>% 
  bind_cols(as_tibble(banana_autoarima_forecast$upper)) %>% 
  rename(`Upper 80%` = `80%`, `Upper 95%` = `95%`) %>% 
  mutate(Type = "Forecast") %>% 
  bind_rows()
```

```{r AA Log Return TS, fig.height=3, fig.width=7}
banana_forecast_logreturn <- 
  banana_price %>% 
  select(
    Date
    ,Mean = `Log Return`
  ) %>% 
  mutate(
    `Lower 80%` = Mean
    ,`Lower 95%` = Mean
    ,`Upper 80%` = Mean
    ,`Upper 95%` = Mean
    ,Type = "Actual"
  ) 

banana_autoarima_logreturn_plot <- 
  banana_forecast_logreturn %>% 
  bind_rows(banana_autoarima_forecast) %>% 
  bind_rows(
    banana_forecast_logreturn %>% 
      filter(Date == max(Date)) %>% 
      mutate(Type = "Forecast")
  ) %>% 
  mutate(Model = "Based on the Auto-ARIMA Model for Log Returns") %>% 
  ggplot(
    aes(
      x = Date
      ,y = Mean
      ,colour = Type
    )
  ) +
  geom_ribbon(
    aes(ymin = `Lower 95%`, ymax = `Upper 95%`)
    ,fill = "grey60"
    ,alpha = 0.4
  ) +
  geom_line() +
  facet_wrap(~ Model) +
  scale_colour_brewer(
    type = "seq"
    ,palette = "Set2"
  ) +
  theme(axis.title.y = element_blank())
```

```{r AA Forecast Price, fig.height=3, fig.width=7}
banana_forecast_price <- 
  banana_price %>% 
  select(
    Date
    ,Mean = Price
  ) %>% 
  mutate(
    `Lower 80%` = Mean
    ,`Lower 95%` = Mean
    ,`Upper 80%` = Mean
    ,`Upper 95%` = Mean
    ,Type = "Actual"
  )

banana_forecast_price <-
  banana_forecast_price %>% 
  bind_rows(banana_forecast_price) %>% 
  bind_rows(
    banana_forecast_price %>% 
      filter(Date == max(Date)) %>% 
      mutate(Type = "Forecast")
  )


for (t in 1:nrow(banana_autoarima_forecast)){
  banana_forecast_price <- 
    log_ret_to_price(
      historical_data = banana_forecast_price
      ,forecast_data = banana_autoarima_forecast
    )
}
```

```{r AA Price TS, fig.height=3, fig.width=7}
banana_autoarima_price_plot <- 
  banana_forecast_price %>% 
  mutate(Model = "Based on the Auto-ARIMA Model for Log Returns") %>% 
  ggplot(
    aes(
      x = Date
      ,y = Mean
      ,colour = Type
    )
  ) +
  geom_ribbon(
    aes(ymin = `Lower 95%`, ymax = `Upper 95%`)
    ,fill = "grey60"
    ,alpha = 0.4
  ) +
  geom_line() +
  facet_wrap(~ Model) +
  scale_colour_brewer(
    type = "seq"
    ,palette = "Set2"
  ) +
  theme(axis.title.y = element_blank())
```

```{r AIC Forecast Log Return, fig.height=3,fig.width=7}
banana_aic_forecast <- 
  banana_aic_arima$`1` %>% 
  forecast(num_periods_ahead) 

banana_aic_forecast <-
  tibble(
    Date = max(banana_price$Date) + months(1:num_periods_ahead)
    ,Mean = as.numeric(banana_aic_forecast$mean)
  ) %>% 
  bind_cols(as_tibble(banana_aic_forecast$lower)) %>% 
  rename(`Lower 80%` = `80%`, `Lower 95%` = `95%`) %>% 
  bind_cols(as_tibble(banana_aic_forecast$upper)) %>% 
  rename(`Upper 80%` = `80%`, `Upper 95%` = `95%`) %>% 
  mutate(Type = "Forecast") %>% 
  bind_rows()
```

```{r AIC Log Return TS, fig.height=3,fig.width=7}
banana_aic_logreturn_plot <- 
  banana_forecast_logreturn %>% 
  bind_rows(banana_aic_forecast) %>% 
  bind_rows(
    banana_forecast_logreturn %>% 
      filter(Date == max(Date)) %>% 
      mutate(Type = "Forecast")
  ) %>% 
  mutate(Model = "Based on the Minimum AIC Model for Log Returns") %>% 
  ggplot(
    aes(
      x = Date
      ,y = Mean
      ,colour = Type
    )
  ) +
  geom_ribbon(
    aes(ymin = `Lower 95%`, ymax = `Upper 95%`)
    ,fill = "grey60"
    ,alpha = 0.4
  ) +
  geom_line() +
  facet_wrap(~ Model) +
  scale_colour_brewer(
    type = "seq"
    ,palette = "Set2"
  ) +
  theme(axis.title.y = element_blank())
```

```{r AIC Forecast Price, fig.height=3, fig.width=7}
banana_forecast_price <- 
  banana_price %>% 
  select(
    Date
    ,Mean = Price
  ) %>% 
  mutate(
    `Lower 80%` = Mean
    ,`Lower 95%` = Mean
    ,`Upper 80%` = Mean
    ,`Upper 95%` = Mean
    ,Type = "Actual"
  )

banana_forecast_price <-
  banana_forecast_price %>% 
  bind_rows(banana_forecast_price) %>% 
  bind_rows(
    banana_forecast_price %>% 
      filter(Date == max(Date)) %>% 
      mutate(Type = "Forecast")
  )


for (t in 1:nrow(banana_aic_forecast)){
  banana_forecast_price <- 
    log_ret_to_price(
      historical_data = banana_forecast_price
      ,forecast_data = banana_aic_forecast
    )
}
```

```{r AIC Price TS, fig.height=3, fig.width=7}
banana_aic_price_plot <- 
  banana_forecast_price %>% 
  mutate(Model = "Based on the Minimum AIC Model for Log Returns") %>% 
  ggplot(
    aes(
      x = Date
      ,y = Mean
      ,colour = Type
    )
  ) +
  geom_ribbon(
    aes(ymin = `Lower 95%`, ymax = `Upper 95%`)
    ,fill = "grey60"
    ,alpha = 0.4
  ) +
  geom_line() +
  facet_wrap(~ Model) +
  scale_colour_brewer(
    type = "seq"
    ,palette = "Set2"
  ) +
  theme(axis.title.y = element_blank())
```

```{r BIC Forecast Log Return}
banana_bic_forecast <- 
  banana_bic_arima$`2` %>% 
  forecast(num_periods_ahead) 

banana_bic_forecast <-
  tibble(
    Date = max(banana_price$Date) + months(1:num_periods_ahead)
    ,Mean = as.numeric(banana_bic_forecast$mean)
  ) %>% 
  bind_cols(as_tibble(banana_bic_forecast$lower)) %>% 
  rename(`Lower 80%` = `80%`, `Lower 95%` = `95%`) %>% 
  bind_cols(as_tibble(banana_bic_forecast$upper)) %>% 
  rename(`Upper 80%` = `80%`, `Upper 95%` = `95%`) %>% 
  mutate(Type = "Forecast") %>% 
  bind_rows()
```

```{r BIC Log Return TS, fig.height=3, fig.width=7}
banana_bic_logreturn_plot <-
  banana_forecast_logreturn %>% 
  bind_rows(banana_bic_forecast) %>% 
  bind_rows(
    banana_forecast_logreturn %>% 
      filter(Date == max(Date)) %>% 
      mutate(Type = "Forecast")
  ) %>% 
  mutate(Model = "Based on the Minimum BIC Model for Log Returns") %>% 
  ggplot(
    aes(
      x = Date
      ,y = Mean
      ,colour = Type
    )
  ) +
  geom_ribbon(
    aes(ymin = `Lower 95%`, ymax = `Upper 95%`)
    ,fill = "grey60"
    ,alpha = 0.4
  ) +
  geom_line() +
  facet_wrap(~ Model) +
  scale_colour_brewer(
    type = "seq"
    ,palette = "Set2"
  ) +
  theme(axis.title.y = element_blank())
```

```{r BIC Forecast Price}
banana_forecast_price <- 
  banana_price %>% 
  select(
    Date
    ,Mean = Price
  ) %>% 
  mutate(
    `Lower 80%` = Mean
    ,`Lower 95%` = Mean
    ,`Upper 80%` = Mean
    ,`Upper 95%` = Mean
    ,Type = "Actual"
  )

banana_forecast_price <-
  banana_forecast_price %>% 
  bind_rows(banana_forecast_price) %>% 
  bind_rows(
    banana_forecast_price %>% 
      filter(Date == max(Date)) %>% 
      mutate(Type = "Forecast")
  )


for (t in 1:nrow(banana_bic_forecast)){
  banana_forecast_price <- 
    log_ret_to_price(
      historical_data = banana_forecast_price
      ,forecast_data = banana_bic_forecast
    )
}
```

```{r BIC Price TS, fig.height=3, fig.width=7}
banana_bic_price_plot <-
  banana_forecast_price %>% 
  mutate(Model = "Based on the Minimum BIC Model for Log Returns") %>%  
  ggplot(
    aes(
      x = Date
      ,y = Mean
      ,colour = Type
    )
  ) +
  geom_ribbon(
    aes(ymin = `Lower 95%`, ymax = `Upper 95%`)
    ,fill = "grey60"
    ,alpha = 0.4
  ) +
  geom_line() +
  facet_wrap(~ Model) +
  scale_colour_brewer(
    type = "seq"
    ,palette = "Set2"
  ) +
  theme(axis.title.y = element_blank())
```


## Log Return Comparison
```{r Comparison Log Return}
grid.arrange(
  banana_autoarima_logreturn_plot
  ,banana_aic_logreturn_plot
  ,banana_bic_logreturn_plot
  ,ncol = 1
  ,top = "Forecasted Banana Log Return"
  ,left = "Log Return"
)
```

## Price Comparison
```{r Comparison Price}
grid.arrange(
  banana_autoarima_price_plot
  ,banana_aic_price_plot
  ,banana_bic_price_plot
  ,ncol = 1
  ,top = "Forecasted Banana Price"
  ,left = "Price (USD/Metric Ton)"
)
```


## Accuracy Comparison
```{r Comparison Accuracy}
bind_cols(
  Model = 
    c(
      "Auto-ARIMA"
      ,"Minimum AIC"
      ,"Minimum BIC"
    )
  ,bind_rows(
    banana_log_ret_autoarima_01 %>% 
      accuracy() %>% 
      as_tibble()
    ,banana_aic_arima$`1` %>% 
      accuracy() %>% 
      as_tibble()
    ,banana_bic_arima$`1` %>% 
      accuracy() %>% 
      as_tibble()
  )
) %>% 
  select(-MPE, -MAPE) %>% 
  kable(caption = "Comparison of Model Accuracy")
```

# Model Diagnostics
```{r Model Diag, include=FALSE}
banana_aic_arima_d <- 
  banana_price$`Log Return` %>% 
  Arima(
    order = banana_min_aic[1, 1:3] %>% t()
    ,seasonal = banana_min_aic[1, 4:6] %>% t()
    ,include.drift = TRUE
  )
banana_aic_arima_s <-
  banana_price$`Log Return` %>% 
  Arima(
    order = banana_min_aic[1, 1:3] %>% t
    ,seasonal = banana_min_aic[1, 4:6] %>% t() %>% c(3)
  )
```

## White Noise
```{r White Noise}
tsdiag(banana_aic_arima$`1`)
# tsdisplay(residuals(banana_aic_arima$`1`), lag.max=23, main='Figure 23: ARIMA(1,0,1) Model Residuals')
```


## Normality of Residuals
```{r}
shapiro.test(residuals(banana_aic_arima$`1`))
jarque.bera.test(residuals(banana_aic_arima$`1`)) 
```


## Constant Variance
```{r}
plot(residuals(banana_aic_arima$`1`)[1:240], residuals(banana_aic_arima$`1`)[2:241], pch=20,main="Scatterplot of Residual with Lag 1") 
abline(h=0,lty=2)
```


# Results
## Model
$$
\phi(B)\Phi(B^s)Y_t= \theta(B)\Theta(B^s)Z_t
$$
$$
\phi(z) = 1-\phi_1(z)-...-\phi_pz^p,\Phi(z) = 1-\Phi_1z-...-\Phi_pz^p
$$

ARIMA (1,0,1)
$$
\phi(z) = 1-0.7573z
$$ 
$$
\theta(z) = 1-0.9568z
$$


# Appendix

## References
[1] Federal Reserve Bank of St.Louis (FRED). (2017, July 12). Global price of Bananas. Retrieved April 09, 2018, from https://fred.stlouisfed.org/series/PBANSOPUSDM

[2] Britannica, T. E. (2018, March 15). Banana. Retrieved April 09, 2018, from http://www.britannica.com/plant/banana-plant

[3] Banana Link. (n.d.). Retrieved April 09, 2018, from http://www.bananalink.org.uk/
